Tulai

### 2023-09-07 Tulai.qmd
test

---
title: "Tulai"
author: "S.SH"
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

@@

```{r}
#exploring Data
# barplot of lithic frequency based on layers

@@

tl2 <- tl1 %>%
  mutate(
    Blank = case_when(
      Breakage %in% c("0", "1", NA_character_) ~ Blank,  # If Breakage is 0, 1, or NA, keep Blank unchanged
      TRUE ~ Breakage  # If Breakage is other than 0, 1, or NA, transfer its value to Blank
    )
  )

@@

```{r}
# First convert all columns to character type
tl4 <- tl4 %>%
  mutate(across(everything(), as.character))

@@

```{r}
# Update the 'typology-tool-core' column based on the condition in 'Retouch'
tl_final <- tl_final %>% 
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",

@@

# Creating tooltypegroup based on the given criteria

@@

```{r}
tl_final <- tl_final %>%
  mutate(
    `core-fragment` = if_else(
      !is.na(`core-typology`),  # If core-typology is not NA
      NA_character_,            # Set core-fragment to NA
      `core-fragment`           # Otherwise, keep the original value
    )
  )

# Show updated data frame
print(tl_final)

@@

                   #DO NOT RUN#############

tl_final <- tl_final %>%
  mutate(
    `core-fragments` = str_to_lower(`core-fragment`),
    `core-fragments` = str_trim(`core-fragment`),
    `core-fragments` = str_replace_all(`core-fragment`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(`core-fragment` = case_when(

            `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece"

            `core-fragment` == "crested bladelet" ~ "crested bladelet",

            `core-fragment` == "crested bladelet" ~ "crested bladelet",

            `core-fragment` == "primaryflake" ~ "primary flake",
            `core-fragment` == "primary flake" ~ "primary flake",

            `core-fragment` == "primary blade" ~ "primary blade",

            `core-fragment` == "rejuvenationpiece-side" ~ "rejuvenation piece-side",
            `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece-side",

            `core-fragment` == "multidirectional core fragment" ~ "shapeless",
            `core-fragment` == "multidirectionalcore fragment" ~ "shapeless",
            `core-fragment` == "multidirectionalbladelet core" ~ "shapeless",

            `core-fragment` == ("bipolar-percussion") ~ "cylinder/prismatic-percussion",

            `core-fragment` == "core platform rejuvention" ~ "core tablet",
            `core-fragment` == "bipolar" ~ "cylinder/prismatic",
             `core-fragment` == "coretablet" ~ "core tablet",
            `core-fragment` == "crestedbladelet" ~ "crested piece",
            `core-fragment` == "coreon flake" ~ "core on flake",
            `core-fragment` == "corefragment-pyramid" ~ "pyramid",
            `core-fragment` == "corepreparation piece-platform" ~ "core tablet",

    TRUE ~ `core-fragment`
  ))

unique(`tl_final`$`core-fragment`)
```

@@

# Process the tl_final_area dataframe using pipe (%>%)
tl_final <- tl_final %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(corefrag = case_when(
    `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
    `core-typology` == "core fragment" ~ "core fragment",
    TRUE ~ corefrag
  )) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology` but keep the column
  mutate(`core-typology` = if_else(`core-typology` %in% c("rejuvenation piece", "core fragment"), NA_character_, `core-typology`))

@@

# Show the updated data
print(tl_final)
unique(tl_final$`core-fragment`)
unique(tl_final$`blank`)

@@

Summary tables: 
1.AREA
```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) & corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# Show summary table
print(summary_table_area)

@@

```{r}
# Create a summary table based on the 'depth' column
summary_table_depth <- tl_final %>%
  group_by(depth) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Create a summary table based on the 'depth', 'area', and 'layer' columns
summary_table_depth_area_layer <- tl_final %>%
  group_by(depth, area, layer, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Load dplyr library
library(dplyr)

@@

```{r}
#Generate summary table-area

@@

# Create a row that contains the column totals, and append it to the table
total_row <- summary_table_tool_area %>%
  select(-tooltypegroup) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  mutate(tooltypegroup = "Total")

# Combine the total_row with the original table
summary_table_tool_area <- bind_rows(summary_table_tool_area, total_row)

# View the summary table
print(summary_table_tool_area)

@@

kable(summary_table_tool_area, 
      col.names = c("Tools", "D1", "TA", "TP1", "Total"),
      caption = "Table: Tool Typology in three different main areas.", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  column_spec(1, extra_css = "text-align:left;") %>%  # Make the first column left-aligned
  row_spec(0, bold = TRUE, font_size = 14)

@@

volume for each area-layer: 
Ta-

```{r}
# Define the volumes for each area based on the given information
volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375) # in m^3
)

@@

consistency in dimention:
```{r}
#|label: fig_2
#|fig-cap: "Boxplot for each Dimention"
#|fig-height: 5

@@

1.   (formal + informal) to debitages
2.  formal  vs. informal
3.  shine/ sickle  as a  used for processing plants

@@

```{r}
# Load necessary libraries
# Import required packages
library(ggplot2)
library(ggpmisc)

# Create the data frame based on the example table
summary_table2_area <- data.frame(
  area = c("D1_a", "D1_c", "TA_d", "TA_e", "TP1_c"),
  lithic_volumetric_density = c(51.43, 111.11, 45.13, 100.08, 212.50),
  retouch_frequency = c(0.05, 0.09, 0.07, 0.07, 0.10)
)

# Generate the regression plot
p <- ggplot(summary_table2_area, aes(x = lithic_volumetric_density, y = retouch_frequency)) +

  # Titles and labels
  labs(title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
       x = "Lithic Volumetric Density",
       y = "Retouch Frequency")

```{r}
#plot_blank_type
# Libraries
library(dplyr)
library(ggplot2)
library(ggbeeswarm)

# Assuming your data is stored in the variable tl4_tool_Blank
# Filter the data to keep only bladelet, blade, and flake in the Blank column
tl_final_aread <- tl_final_area %>%
  filter(blank %in% c("bladelet", "blade", "flake"))  %>%

# Group the data by Blank2 and calculate the counts for each combination of Utilization, shine2, and Retouch

# Print the table using kable() and kableExtra for formatting
table_kable <- kable(table_summary, format = "html", caption = "Summary Table", 
                     col.names = c("Blank2", "Utilization", "shine2", "Retouch", "Count")) %>%
  kable_styling(bootstrap_options = "striped")

# Print the table
print(table_kable)

# Plot
# Pivot data to longer format
tl4_tool_Blank_long <- 
  tl4_tool_Blank_selected %>% 
  pivot_longer(cols = c(Utilization, Retouch, shine2), 
               names_to = "Blank2", 
               values_to = "Value")

# Create boxplot
ggplot(tl4_tool_Blank_long, aes(x = Dimension, y = Value)) +
  geom_boxplot() +
  facet_wrap(~, scales = "free") +
  labs(x = "Blank", 
       y = "Frequency") +
  theme_minimal()

# Print the plot
print(tl4_plots)

# Print the combined plot
print(plot_combined)

WABI: 
```{r}
# regression plot-layer
library(ggplot2)
library(ggpmisc)

tl_final_area %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d")) %>%
  filter(!is.na(corefrag)) %>% 
  select(layer, corefrag) %>%
  ggplot() +
  aes(x = corefrag) +
  geom_bar() +  # Use geom_bar for categorical data
  labs(x = "Core and Core fragments", y = "Frequency") +
  facet_wrap(~layer, ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```{r}
# Calculating Blank Frequency
blank_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(blank)) %>%
  group_by(layer) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca,
             axes = 2,   # Selecting the second axis
             choice = "var") # Choosing to show contributions of variables

```{r}
# Load necessary libraries
library(FactoMineR)
library(factoextra)

```{r}
#feature frequency: 
d1     1 
ta    1
tp1 	0)

```{r}
#estimated area
A→  52 m3
TP → 5.76 m 3
D1 → 2.1375 m3
```

```{r}
# Extract the loadings (coordinates of variables)
loadings <- res_pca_area$var$coord

# Extract the eigenvalues
eigenvalues <- res_pca_area$eig

# Calculate the contribution for Dim.1
contrib_dim1 <- (loadings[,1]^2 / eigenvalues[1, 1]) * 100

# Calculate the contribution for Dim.2
contrib_dim2 <- (loadings[,2]^2 / eigenvalues[2, 1]) * 100

# Create a data frame to hold the contributions
contrib_df <- data.frame(
  Variable = rownames(loadings),
  Contrib_Dim1 = contrib_dim1,
  Contrib_Dim2 = contrib_dim2
)

# Show the contributions
print(contrib_df)

# Generate the scree plot
fviz_eig(res_pca_area)

```{r}
# Extract individual coordinates
individual_coordinates <- res_pca_area$ind$coord

# Convert to a data frame for better visibility and add the 'area' column back for context
individual_coordinates_df <- as.data.frame(individual_coordinates)
individual_coordinates_df$area <- unique(area_data$area)

# Show the data frame
print(individual_coordinates_df)

# Create the table using gt
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```{r}
# Count the number of artifacts by area and store it in a new data frame
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Calculating retouch Frequency

# Calculating Blank Frequency
blank_freq <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(blank)) %>%
  group_by(area) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

```{r}
# Load necessary libraries
library(FactoMineR)
library(factoextra)

# Create the table using gt
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```{r}
# Count the number of artifacts by area and store it in a new data frame
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

```{r}
# Define the volumes for each area based on the given information
volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375)
)

>>>>

---
title: "Tulai"
author: ""
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

@@

@@

tl2 <- tl1 %>%
  mutate(
    Blank = case_when(
      Breakage %in% c("0", "1", NA_character_) ~ Blank,

      TRUE ~ Breakage  
    )
  )

@@

```{r}
# convert all columns to character type (I DONT KNOW WHY I DID THIS!)
tl4 <- tl4 %>%
  mutate(across(everything(), as.character))

@@

```{r}
# Update the 'typology-tool-core' column based on 'Retouch'
tl_final <- tl_final %>% 
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",

@@

# Creating tooltypegroup

@@

```{r}
tl_final <- tl_final %>%
  mutate(
    `core-fragment` = if_else(
      !is.na(`core-typology`),  
      NA_character_,            
      `core-fragment`           
    )
  )

@@

@@

tl_final <- tl_final %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(corefrag = case_when(
    `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
    `core-typology` == "core fragment" ~ "core fragment",
    TRUE ~ corefrag
  )) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology` 
  mutate(`core-typology` = if_else(`core-typology` %in% c("rejuvenation piece", "core fragment"), NA_character_, `core-typology`))

@@

print(tl_final)
unique(tl_final$`core-fragment`)
unique(tl_final$`blank`)

@@

Summary tables: 1.AREA

```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) & corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

print(summary_table_area)

@@

```{r}
# Create a summary table based on the 'depth' column
summary_table_depth <- tl_final %>%
  group_by(depth) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Create a summary table based on the 'depth', 'area', and 'layer' columns
summary_table_depth_area_layer <- tl_final %>%
  group_by(depth, area, layer, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

@@

```{r}
#summary table-area

@@

total_row <- summary_table_tool_area %>%
  select(-tooltypegroup) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  mutate(tooltypegroup = "Total")

summary_table_tool_area <- bind_rows(summary_table_tool_area, total_row)

print(summary_table_tool_area)

@@

kable(summary_table_tool_area, 
      col.names = c("Tools", "D1", "TA", "TP1", "Total"),
      caption = "Table: Tool Typology in three different main areas.", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  column_spec(1, extra_css = "text-align:left;") %>%  
  row_spec(0, bold = TRUE, font_size = 14)

@@

volume for each area:

```{r}
# Define the volumes for each area 
volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375) # in m^3
)

@@

library(tidyverse)
library(ggpmisc)

# Create the data frame
summary_table2_area <- tibble(
  area = c("D1_a", "D1_c", "TA_d", "TA_e", "TP1_c"),
  lithic_volumetric_density = c(51.43, 111.11, 45.13, 100.08, 212.50),
  retouch_frequency = c(0.05, 0.09, 0.07, 0.07, 0.10)
)

# Create the ggplot
p <- ggplot(summary_table2_area, 
            aes(x = lithic_volumetric_density,
                                     y = retouch_frequency)) +

  # Add points
  geom_point(color = 
               "black", 
             size = 2) +

  # line with log scale
  stat_smooth(method = "lm", 
              formula = y ~ log(x), 
              se = TRUE, aes(group = 1), 
              color = "blue", 
              size = 0.5) +

  # Add text labels for each point
  geom_text(aes(label = area), 
            vjust = -1, 
            hjust = -0.5, 
            size = 2.5) +

  # Add regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), stat(rr.label), stat(p.value.label), sep = "~~~")),
               formula = y ~ log(x), 
               parse = TRUE, 
               size = 2) +

  # Add titles and labels
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  scale_x_log10() +

  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

library(googlesheets4)

data_sheet <- "https://docs.google.com/spreadsheets/d/16yLb-RyIQ7x2jFEH_Q7OZYTwlEAXJgFu/edit#gid=2014421097"

sum_depth_sheet <- read_sheet(data_sheet, sheet = "sum_depth")

```{r}
#plot_blank_type

library(dplyr)
library(ggplot2)
library(ggbeeswarm)

#table

tl_final_aread <- tl_final_area %>%
  filter(blank %in% c("bladelet", "blade", "flake"))  %>%

table_kable <- kable(table_summary, format = "html", caption = "Summary Table", 
                     col.names = c("Blank2", "Utilization", "shine2", "Retouch", "Count")) %>%
  kable_styling(bootstrap_options = "striped")

print(table_kable)

tl4_tool_Blank_long <- 
  tl4_tool_Blank_selected %>% 
  pivot_longer(cols = c(Utilization, Retouch, shine2), 
               names_to = "Blank2", 
               values_to = "Value")

ggplot(tl4_tool_Blank_long, aes(x = Dimension, y = Value)) +
  geom_boxplot() +
  facet_wrap(~, scales = "free") +
  labs(x = "Blank", 
       y = "Frequency") +
  theme_minimal()

print(tl4_plots)

print(plot_combined)

WABI:

```{r}
# regression plot-layer
library(ggplot2)
library(ggpmisc)

tl_final_area %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d")) %>%
  filter(!is.na(corefrag)) %>% 
  select(layer, corefrag) %>%
  ggplot() +
  aes(x = corefrag) +
  geom_bar() +  
  labs(x = "Core and Core fragments", y = "Frequency") +
  facet_wrap(~layer, ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```{r}
# Blank Frequency
blank_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(blank)) %>%
  group_by(layer) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

feature frequency: d1 1 ta 1 tp1 0)

estimated volumes (area) A→ 52 m3 TP → 5.76 m 3 D1 → 2.1375 m3

loadings <- res_pca_area$var$coord

# 
eigenvalues <- res_pca_area$eig

#  Dim.1
contrib_dim1 <- (loadings[,1]^2 / eigenvalues[1, 1]) * 100

#  Dim.2
contrib_dim2 <- (loadings[,2]^2 / eigenvalues[2, 1]) * 100

contrib_df <- data.frame(
  Variable = rownames(loadings),
  Contrib_Dim1 = contrib_dim1,
  Contrib_Dim2 = contrib_dim2
)

print(contrib_df)

fviz_eig(res_pca_area)

individual_coordinates <- res_pca_area$ind$coord

individual_coordinates_df <- as.data.frame(individual_coordinates)
individual_coordinates_df$area <- unique(area_data$area)

print(individual_coordinates_df)

gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

blank_freq <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(blank)) %>%
  group_by(area) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Create the table 
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```{r}
# Count the number of artifacts by area
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375)
)

consistency in dimention:

```{r}
#|label: fig_2
#|fig-cap: "Boxplot for each Dimention"
#|fig-height: 5

@@

@@
